---
title: "S305N Cell Lines RNAseq Analysis"
author:
  - name: "Emir Turkes and Eliona Tsefou, UK DRI at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: "../s305n-clines-rnaseq.bib"
biblio-style: apalike
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path("..", "results", "s305n_clines_rnaseq_analysis.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [s305n-clines-rnaseq](https://github.com/eturkes/s305n-clines-rnaseq).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.  
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.  
The toggle at the start of the document controls the visibility of all chunks.  

This analysis is a comparison of bulk RNAseq data from human iPSC neuronal cell lines  
They were obtained from Synthego and NCRAD and feature the MAPT S305 mutation.  
We also compare bulk RNAseq data from @bowles_development_2024, which used the same NCRAD lines.

```{r, message = FALSE, warning = FALSE, results = "hide"}
# Copyright (c) Emir Turkes, Eliona Tsefou, UK DRI at UCL
#
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
#
#    http://www.apache.org/licenses/LICENSE-2.0
#
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.
#
# ---- LLVM Exceptions to the Apache 2.0 License ----
#
# As an exception, if, as a result of your compiling your source code, portions
# of this Software are embedded into an Object form of such source code, you
# may redistribute such embedded portions in such Object form without complying
# with the conditions of Sections 4(a), 4(b) and 4(d) of the License.
#
# In addition, if you combine or link compiled forms of this Software with
# software that is licensed under the GPLv2 ("Combined Software") and if a
# court of competent jurisdiction determines that the patent provision (Section
# 3), the indemnity provision (Section 9) or other Section of the License
# conflicts with the conditions of the GPLv2, you may retroactively and
# prospectively choose to deem waived or otherwise exclude such Section(s) of
# the License, but only in their entirety and only with respect to the Combined
# Software.

# Load packages.
# --------------
library(conflicted) # Ensure packages do not have namespace conflicts.
packages <- "edgeR"
lapply(packages, library)
# --------------

# Adds download buttons and horizontal scrolling to DT::datatable.
# ----------------------------------------------------------------
datatable_download <- function(dt) {
  DT::datatable(
    dt,
    list(
      scrollX = TRUE, dom = "Blfrtip",
      buttons = list(
        "copy", "print",
        list(
          extend = "collection", buttons = c("csv", "excel", "pdf"),
          text = "Download"
        )
      )
    ),
    extensions = "Buttons"
  )
}
# ----------------------------------------------------------------

# Same as datatable_download but converts numbers to exponential values.
# ----------------------------------------------------------------------
datatable_download_exp <- function(dt) {
  DT::datatable(
    dt,
    list(
      scrollX = TRUE,
      dom = "Blfrtip",
      buttons = list(
        "copy", "print",
        list(
          extend = "collection", buttons = c("csv", "excel", "pdf"),
          text = "Download"
        )
      ),
      rowCallback = DT::JS(
        "function(row, data) {",
        "for (i = 1; i < data.length; i++) {",
        "if (data[i]>=1000 | data[i]<1000) {",
        "$('td:eq('+i+')', row).html(data[i].toExponential(2));}}}"
      )
    ),
    extensions = "Buttons"
  )
}
# ----------------------------------------------------------------------

# Add paths.
# ----------
data_dir <- file.path("..", "data")
cache_dir <- file.path("..", "cache")
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ----------
```

# Prep Data

```{r}
data <- vector("list", length = 3)

data[[1]] <- read.csv(file.path(data_dir, "Final counts_RNAseq_Synthego.txt"), sep = "\t", row.names = 1)
data[[2]] <- read.csv(file.path(data_dir, "gene_count_NCRAD.txt"), sep = "\t", row.names = 1)

# Data uses ENSEMBL IDs by default but includes gene symbols.
# Use one of the datasets to create a conversion table.
# Then strip extra information from the datasets that are not counts.
# -------------------------------------------------------------------
gene_anno <- data.frame(
  ensembl = rownames(data[[1]]), symbols = data[[1]]$gene_name
)
for (i in 1:2) {
  data[[i]] <- data[[i]][ , seq(which(colnames(data[[i]]) == "gene_name") - 1)]
}
# -------------------------------------------------------------------

# When stripping bowles data to plain ENSEMBL IDs there are duplicates.
# We collapse these duplicates to the mean of each set per sample.
# Since this process is somewhat slow, we cache the cleaned object.
# ---------------------------------------------------------------------
cache <- file.path(cache_dir, "bowles.qs")
if (file.exists(cache)) {
  data[[3]] <- qs::qread(cache)
} else {
  data[[3]] <- read.csv(file.path(data_dir, "GSE260517_star_gene.txt"), sep = "\t")
  data[[3]]$gid <- sub("\\..*", "", data[[3]]$gid)
  data[[3]] <- aggregate(. ~ gid, data = data[[3]], FUN = mean)
  rownames(data[[3]]) <- data[[3]]$gid
  data[[3]]$gid <- NULL
  qs::qsave(data[[3]], file = cache)
}
# ---------------------------------------------------------------------

# Subset datasets to samples of interest.
# ---------------------------------------
match <- paste(
  c(
    "WT21", "WT28", "D421", "D428",
    "NNBB7_D21", "NNBB7_D28", "HCB8_D21", "HCB8_D28",
    "NWBB7_6W", "NWBB7_8W", "HCB8_6W", "HCB8_8W"
  ),
  collapse = "|"
)
data <- lapply(
  data, function(mat) mat[, grepl(match, colnames(mat)), drop = FALSE]
)
# ---------------------------------------

names(data) <- c("synthego", "ncrad", "bowles")
```

**Preview of Synthego RNAseq Data**

```{r}
datatable_download(data$synthego[1:5, ])
```

**Preview of NCRAD RNAseq Data**

```{r}
datatable_download(data$ncrad[1:5, ])
```

**Preview of NCRAD RNAseq Data from @bowles_development_2024**

```{r}
datatable_download(data$bowles[1:5, ])
```

# Differential Expression

```{r}
# Create a list of metadata that will be used for design matrices.
# ----------------------------------------------------------------
meta <- vector("list", length = 3)

cols <- colnames(data[[1]])
meta[[1]] <- data.frame(
  sample = cols,
  timepoint = factor(sub(".*(21|28).*", "\\1", cols)),
  genotype = factor(sub("^(WT|D4).*", "\\1", cols))
)
cols <- colnames(data[[2]])
meta[[2]] <- data.frame(
  sample = cols,
  timepoint = factor(sub(".*_D(\\d+)$", "\\1", cols)),
  genotype = factor(sub(".*_(NNBB7|HCB8)_D\\d+$", "\\1", cols))
)
cols <- colnames(data[[3]])
meta[[3]] <- data.frame(
  sample = cols,
  timepoint = factor(sub(".*_(6W|8W)$", "\\1", cols)),
  genotype = factor(sub(".*_(HCB8|NWBB7)_[68]W$", "\\1", cols))
)
# ----------------------------------------------------------------

# Run limma-trend pipeline on each dataset.
# For now the third dataset is excluded for having too few samples.
# -----------------------------------------------------------------
cache <- file.path(cache_dir, "results.qs")
if (file.exists(cache)) {
  results <- qs::qread(cache)
} else {

  results <- vector("list", length = 2)
  names(results) <- c("synthego", "ncrad")
  for (i in seq_along(results)) {

    dge_orig <- DGEList(data[[i]])
    keep <- filterByExpr(dge_orig, group = meta[[i]]$genotype)
    dge_orig <- dge_orig[keep, ]
    dge_orig <- calcNormFactors(dge_orig)
    dge <- DGEList(data[[i]])
    dge$samples$norm.factors <- dge_orig$samples$norm.factors
    keep <- filterByExpr(dge_orig, group = meta[[i]]$genotype)
    dge <- dge[keep, , keep.lib.sizes = FALSE]
    dge <- calcNormFactors(dge)
    dge <- cpm(dge, log = TRUE)

    group <- factor(paste(meta[[i]]$genotype, meta[[i]]$timepoint, sep = "_"))
    design <- model.matrix(~ 0 + group)
    colnames(design) <- levels(group)

    if (i == 1) {
      cont_mat <- makeContrasts(
        D4_vs_WT_21 = D4_21 - WT_21,
        D4_vs_WT_28 = D4_28 - WT_28,
        levels = design
      )
    } else if (i == 2) {
      cont_mat <- makeContrasts(
        HCB8_vs_NNBB7_21 = HCB8_21 - NNBB7_21,
        HCB8_vs_NNBB7_28 = HCB8_28 - NNBB7_28,
        levels = design
      )
    } else if (i == 3) {
      cont_mat <- makeContrasts(
        HCB8_vs_NWBB7_6W = HCB8_6W - NWBB7_6W,
        HCB8_vs_NWBB7_8W = HCB8_8W - NWBB7_8W,
        levels = design
      )
    }

    fit <- lmFit(dge, design)
    fit <- contrasts.fit(fit, cont_mat)
    fit <- suppressWarnings(eBayes(fit, trend = TRUE, robust = TRUE))

    tests <- decideTests(fit)
    tc <- textConnection("tmp", open = "w")
    write.fit(fit, tests, file = tc, adjust = "BH")
    close(tc)
    results[[i]] <- read.delim(text = tmp)

    rownames(results[[i]]) <- results[[i]]$X
    results[[i]] <- results[[i]][ , 2:ncol(results[[i]])]
    gene_anno_sub <- gene_anno[gene_anno$ensembl %in% rownames(results[[i]]), ]
    gene_anno_sub <- gene_anno_sub[match(rownames(results[[i]]), gene_anno_sub$ensembl), ]
    results[[i]]$Gene <- gene_anno_sub$symbol
    results[[i]] <- results[[i]][ , c(ncol(results[[i]]), seq(ncol(results[[i]]) - 1))]
    results[[i]] <- results[[i]][order(results[[i]]$F.p.value), ]
  }
  qs::qsave(results, file = cache)
}
# -----------------------------------------------------------------
```

**Full Synthego DEG Results**

```{r, warning = FALSE}
datatable_download_exp(results$synthego)
```

**Full NCRAD DEG Results**

```{r, warning = FALSE}
datatable_download_exp(results$ncrad)
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
